Multisensor Architecture for an Intersection Management System

### 2018-08-17 ch1-introduction.tex
Final draft
Corrections and figures updated.

Chapter 3 presents the result of three example implementations using the architecture: Single-laser case, multiple laser case and multiple laser and camera case. These test cases are evaluated and compared, both in application performance, using vehicle count as evaluation metric, and in computational resources usage.

>>>>

Chapter 3 presents the result of two example implementations using the architecture: Only camera case and multiple laser and camera case. For the later, 4 different fusion strategies are evaluated. These test cases are implemented and compared using traffic flow status as comparison indicator.

### 2018-08-17 ch3-arch_sys.tex
Corrections and figures updated.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.5]{fig/3/data_model.pdf}
\caption{UML diagram of defined entities and their relationships}
\label{data_model}
\end{figure}

@@

\begin{table}[ht!]
\footnotesize
\centering
\begin{tabular}{|c | p{8cm}|}
\hline
\textbf{Name} & \textbf{Description} \\
\hline
laser\_pol\_msg & Contains a timestamp, an array of N angles and and array of N measurements \\
\hline
laser\_cart\_msg & Contains a timestamp, an array of N (x, y) coordinates \\
\hline
laser\_cfg & Contains position and orientation information, also has a background model for the laser sensor \\
\hline
occgrid\_msg & Contains a timestamp and an occupancy grid in the form of an MxN array \\
\hline
occgrid\_cfg & Contains occupancy grid configuration for the scene \\
\hline
clustering\_msg & Contains a set of clusters, each of them with an ID and a set of points \\
\hline
clustering\_cfg & Contains configuration about clustering algorithm and parameters \\
\hline
\end{tabular}
\caption{Description of diferent types of messages and parameters used}
\label{desc_map}
\end{table}

\todo{Finish table}

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/clusters_veh_counter.pdf}
\caption{Vehicle\_counter}
\label{vehicle_counter}
\end{figure}

\todo{CHANGE FIGURE}

This block accepts different types of inputs, cluster\_msg and occgrid\_msg. It process the incoming data, and generate bounding boxes based on the clusters or based on the occupancy grid, according to the input. Having these bounding boxes, which represent detected vehicles, scene configuration data is loaded and the geometric parameters of the legs and lanes are used to generate virtual sensors at the entry and exit points of the intersection and check the overlaping of detected vehicles with aforemetioned virtual sensors. The output generated by this block is the occupancy percentage of the virtual sensor, represented as a message of type vsens\_occ\_msg.
%vehicle counter $v_{cnt}$ is incremented by one when a transition from occupied state to empty state is detected in a new frame.

@@

\item[virtual\_sensor\_merge] \hfill
\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/veh_cnt_merge.pdf}
\caption{virtual\_sensor\_merge}
\label{virtual_sensor_merge}
\end{figure}

\todo{CHANGE FIGURE}

This block is intended to take as inputs different messages of type vsens\_occ\_msg and merging them to increase the reliabilty of detection. First, a moving average is calculated over each input to reduce the effect of false positives or false negatives in a single frame. After this, the inputs are merged by a defined method like average, weighted average, maximum or minimum. Then, a new vsens\_occ\_msg message is dispatched as output.

@@

\item[virtual\_sensor\_to\_flow\_status] \hfill
\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/veh_cnt_merge.pdf}
\caption{virtual\_sensor\_to\_flow\_statu}
\label{virtual_sensor_to_flow_status}
\end{figure}

\todo{CHANGE FIGURE}

When a message of type vsens\_occ\_msg arrives, it is evaluated using a threshold $vo_{th}$, if the occupancy percentage of a virtual sensor is greater than $vo_{th}$, it is considered that there is a vehicle over the sensor in that frame, this is defined as occupied state. This detection is made in a frame basis, thus, a vehicle counter $v_{cnt}$ is incremented by one when a transition from occupied state to empty state is detected in a new frame.

Then, a time interval $t_{c}$ is defined for calculating the flow rate $v_{r}$. Defined as the number of transitions during that interval divided by $t_{c}$. Then, if $v_{r}$ is above a flow threshold $f_{th}$, it is considered that the traffic is flowing. Otherwise, traffic is considered stopped.

@@

The objective of this block is to allow the visualisation of the data in any of the processing stages. It has the option to display and XY coordinate system, an occupancy grid or a map of the intersection overlaped with detections of metrics.

@@

\subsection{Conclusions}

>>>>

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.75]{fig/3/data_modelA.pdf}
\caption{UML diagram of defined entities and their relationships}
\label{data_model}
\end{figure}

@@

\begin{table}[ht!]
\footnotesize
\centering
\begin{tabular}{|c | p{8cm}|}
\hline
\textbf{Name} & \textbf{Description} \\
\hline
laser\_pol\_msg & Contains a timestamp, an array of N angles and and array of N measurements \\
\hline
laser\_cart\_msg & Contains a timestamp, an array of N (x, y) coordinates \\
\hline
laser\_cfg & Contains position and orientation information, also has a background model for the laser sensor \\
\hline
occgrid\_msg & Contains a timestamp and an occupancy grid in the form of an MxN array \\
\hline
occgrid\_cfg & Contains occupancy grid configuration for the scene \\
\hline
clusters\_msg & Contains a set of clusters, each of them with an ID and a set of points \\
\hline

clusters\_cfg & Contains configuration about clustering algorithm and parameters \\
\hline

camera\_msg & Frames generated by camera \\
\hline

camera\_cfg & Information and parameters of the camera sensor \\
\hline

blobs\_msg & Bounding boxes detected as objects within a frame from the camera \\
\hline

blobs\_cfg & Parameters used by the image detection process\\
\hline

vs\_occ\_msg & Indicates the occupancy level of a virtual sensor, representing the entrace to or exit from the intersection\\
\hline

vs\_merge\_cfg & Parameters for configuring the merging process of different vs\_occ\_ms messages.\\
\hline

flow\_rate\_msg & Indicates the flow rate entering or exiting the intersection\\
\hline

flow\_cfg & Parameters for estimation of the flow rate, flow status and flow merging configuration\\
\hline

flow\_status\_msg & Indicates the flow status of an entrance or exit at intersection. This is a binary value, with value either "Traffic flowing" or "Traffic stopped".\\
\hline

\end{tabular}
\caption{Description of diferent types of messages and parameters used}
\label{desc_map}
\end{table}

%\todo{Finish table}

@@

\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/vs_occ.pdf}
\caption{Virtual\_sensor\_occupancy}
\label{vehicle_counter}
\end{figure}

This block accepts different types of inputs, cluster\_msg and occgrid\_msg. It process the incoming data, and generate bounding boxes based on the clusters or based on the occupancy grid, according to the input. Having these bounding boxes, which represent detected vehicles, scene configuration data is loaded and the geometric parameters of the legs and lanes are used to generate virtual sensors at the entry and exit points of the intersection and check the overlaping of detected vehicles with aforemetioned virtual sensors. The output generated by this block is the occupancy percentage of the virtual sensor, represented as a message of type vs\_occ\_msg.
%vehicle counter $v_{cnt}$ is incremented by one when a transition from occupied state to empty state is detected in a new frame.

@@

\item[virtual\_sensor\_merge] \hfill
\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/vs_merge.pdf}
\caption{virtual\_sensor\_merge}
\label{virtual_sensor_merge}
\end{figure}

This block is intended to take as inputs different messages of type vs\_occ\_msg and merging them to increase the reliabilty of detection. First, a moving average is calculated over each input to reduce the effect of false positives or false negatives in a single frame. After this, the inputs are merged by a defined method like average, weighted average, maximum or minimum. Then, a new vs\_occ\_msg message is dispatched as output.

@@

\item[virtual\_sensor\_to\_flow\_rate] \hfill
\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/vs2flow.pdf}
\caption{virtual\_sensor\_to\_flow\_rate}
\label{virtual_sensor_to_flow_status}
\end{figure}

When a message of type vs\_occ\_msg arrives, it is evaluated using a threshold $vo_{th}$, if the occupancy percentage of a virtual sensor is greater than $vo_{th}$, it is considered that there is a vehicle over the sensor in that frame, this is defined as occupied state. This detection is made in a frame basis, thus, a vehicle counter $v_{cnt}$ is incremented by one when a transition from occupied state to empty state is detected in a new frame.

Then, a time interval $t_{c}$ is defined for calculating the flow rate $v_{r}$. Defined as the number of transitions during that interval divided by $t_{c}$. The output message type is flow\_rate\_msg.

\item[flow\_rate\_to\_flow\_status] \hfill
\begin{figure}[ht!]
\centering
\includegraphics[scale=1]{fig/3/flow2status.pdf}
\caption{flow\_rate\_to\_flow\_status}
\label{flow_rate_to_flow_status}
\end{figure}

This block process one or more inputs of type flow\_rate\_msg. If there are more than one input, it merges them using defined policy, like average or maximum. Then, if the flow rate $v_{r}$ is above a flow threshold $f_{th}$, it is considered that the traffic is flowing. Otherwise, traffic is considered stopped.

@@

The objective of this block is to allow the visualisation of the data in any of the processing stages. It has the option to display and XY coordinate system, an occupancy grid or a map of the intersection overlaped with detections or metrics.

@@

\section{Conclusions}

### 2018-08-17 ch4-results.tex
Corrections and figures updated.

In this chapter is presented the experimental setup used to deploy and test the proposed architecture. First is described the data and hardware and software specifications. Then, the two implemented configurations of the system are presented and finally, results are summarized and analysed.

@@

Situation Assesment &
L & virtual\_sensor\_to\_flow\_status \\
\hline

@@

Each configuration consists of a set of processing blocks (as described in section \ref{proc_blocks}) connected, aiming to take data from  lasers and cameras, to produce an output of higher level. In this case, the binary status of a leg (traffic flow, no traffic flow) is the metric used for evaluation, derived from the vehicle count.

The following table lists all used blocks and assigns an ID for each one. Those IDs are used in the graph description in its own section. Also, bold nodes and connections indicate multiple instance of the same element. The figure \ref{tconf1} shows the  configuration used for processing just video data. Figure \ref{tconf2} shows the configuration when a set of lasers is added to the single-camera configuration.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.7]{fig/4/test_configuration1.pdf}
\caption{Single camera configuration}
\label{tconf1}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.7]{fig/4/test_configuration2.pdf}
\caption{Multiple lasers and camera configuration}
\label{tconf2}
\end{figure}

@@

As stated before, the flow status is used for comparing the system with the ground truth. This status is derived from the flow rate in vehicles per second. If this value is over a threshold, it is considerded to be in "traffic flow" status.

Detection results in three diferent frames are shown in figure \ref{camera_detection}.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.115]{fig/4/camera.jpeg}
\caption{Vehicle detection using camera}
\label{camera_detection}
\end{figure}

\subsection{Case 2: Camera and multiple lasers}

Vehicle detection based on laser scanners is shown in figure \ref{laser_detection}.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.43]{fig/4/laser1.png}
\caption{Vehicle detection using laser scanners}
\label{laser_detection}
\end{figure}

@@

>>>>

In this chapter is presented the experimental setup used to deploy and test the proposed architecture. First is described the data and hardware and software specifications. Then, two implemented configurations of the system are presented. Finally, results are summarized and analysed.

@@

\multirow{2}{*}{Situation Assesment} &
L & virtual\_sensor\_to\_flow\_rate \\
\cline{2-3}
& M & flow\_rate\_to\_flow\_status \\
\hline

@@

Each configuration consists of a set of processing blocks (as described in section \ref{proc_blocks}) connected, aiming to take data from  lasers and cameras, to produce an output of higher level. In this case, the binary status of a leg (traffic flowing, traffic stopped) is the metric used for evaluation, derived from the vehicle count.

The following table lists all used blocks and assigns an ID for each one. Those IDs are used in the graph description in its own section. Also, bold nodes and connections indicate multiple instance of the same element. The figure \ref{tconf1} shows the  configuration used for processing just video data. Figure \ref{tconf2} shows the configuration when a set of lasers is added to the single-camera configuration. The blue line represents the fusion at node K, which is fusion on virtual sensors level. The purple line represents the fusion at flow rate level.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.6]{fig/4/test_configuration1.pdf}
\caption{Single camera configuration}
\label{tconf1}
\end{figure}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.6]{fig/4/test_configuration2.pdf}
\caption{Multiple lasers and camera configuration}
\label{tconf2}
\end{figure}

@@

As stated before, the flow status is used for comparing the system with the ground truth. This status is derived from the flow rate in vehicles per second. If this value is greater than a defined threshold, it is considerded to be in "traffic flowing" status.

As result of the first stages of processing, vehicle detection in three diferent frames are shown in figure \ref{camera_detection}. This is a representation of the output from node H, camera\_blobs.

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.11]{fig/4/camera.jpeg}
\caption{Vehicle detection using camera}
\label{camera_detection}
\end{figure}

In the figure \ref{video_res} is shown the comparison between the flow profile obtained from the system and the generated by the ground truth. There is also remarked one of the intervals where the output is "traffic flowing". After evaluating the performance of the system in a frame basis, a confusion matrix is obtained (table \ref{video_cf}), here $F$ denotes the state "traffic Flowing" and $S$ denotes "traffic Stopped" and we get an accuracy of 91.82\% and a TPR (True Positive Rate) of 55.70\%.

\begin{table}[ht!]
\footnotesize
\centering
\noindent
\renewcommand\arraystretch{1.5}
\setlength\tabcolsep{0pt}
\begin{tabular}{c c c c c}
  \multirow{10}{*}{\rotatebox{90}{\parbox{1.1cm}{\bfseries\centering Actual\\ value}}} & 
    & \multicolumn{2}{c}{\bfseries Outcome} & \\
  & & \bfseries F & \bfseries S & \bfseries total \\
  & F$'$ & \MyBox{694}{} & \MyBox{552}{} & 1246 \\[2.4em]
  & S$'$ & \MyBox{26}{} & \MyBox{5791}{} & 5817 \\
  & total & 720 & 6343 &
\end{tabular}
\caption{Confusion matrix of the system based on video}
\label{video_cf}
\end{table}

\begin{figure}[ht!]
\centering
\includegraphics[scale=0.45]{fig/4/video_res.pdf}
\caption{Results for the single camera implementation in one of the legs of the intersection}
\label{video_res}
\end{figure}

\subsection{Case 2: Camera and multiple lasers}
%{'TN': 5791, 'TP': 694, 'FP': 26, 'FN': 552}
%v30 FPR: 0.0045, F1: 0.7060, NPV: 0.5108, reca: 0.5570, spec: 0.9955, PPV: 0.9639, accu: 0.9182, prec: 0.9639, TPR: 0.5570, prev: 0.1764, 
%{'TN': 5650, 'TP': 23, 'FP': 167, 'FN': 1223}
%l30 FPR: 0.0287, F1: 0.0320, NPV: 0.0156, reca: 0.0185, spec: 0.9713, PPV: 0.1211, accu: 0.8032, prec: 0.1211, TPR: 0.0185, prev: 0.1764, 
%{'TN': 5414, 'TP': 1098, 'FP': 403, 'FN': 148}
%vl_avg_1_30 FPR: 0.0693, F1: 0.7994, NPV: 0.9216, reca: 0.8812, spec: 0.9307, PPV: 0.7315, accu: 0.9220, prec: 0.7315, TPR: 0.8812, prev: 0.1764, 
%{'TN': 5669, 'TP': 792, 'FP': 148, 'FN': 454}
%vl_max_1_30 FPR: 0.0254, F1: 0.7246, NPV: 0.6039, reca: 0.6356, spec: 0.9746, PPV: 0.8426, accu: 0.9148, prec: 0.8426, TPR: 0.6356, prev: 0.1764, 
%{'TN': 5804, 'TP': 161, 'FP': 13, 'FN': 1085}
%vl_f_avg_30 FPR: 0.0022, F1: 0.2268, NPV: 0.1091, reca: 0.1292, spec: 0.9978, PPV: 0.9253, accu: 0.8445, prec: 0.9253, TPR: 0.1292, prev: 0.1764, 
%{'TN': 5624, 'TP': 716, 'FP': 193, 'FN': 530}
%vl_f_max_30 FPR: 0.0332, F1: 0.6645, NPV: 0.5432, reca: 0.5746, spec: 0.9668, PPV: 0.7877, accu: 0.8976, prec: 0.7877, TPR: 0.5746, prev: 0.1764,

%v30         reca: 0.5570, spec: 0.9955, accu: 0.9182, prec: 0.9639, TPR: 0.5570,
%vl_avg_1_30 reca: 0.8812, spec: 0.9307, accu: 0.9220, prec: 0.7315, TPR: 0.8812,
%vl_max_1_30 reca: 0.6356, spec: 0.9746, accu: 0.9148, prec: 0.8426, TPR: 0.6356,
%vl_f_avg_30 reca: 0.1292, spec: 0.9978, accu: 0.8445, prec: 0.9253, TPR: 0.1292,
%vl_f_max_30 reca: 0.5746, spec: 0.9668, accu: 0.8976, prec: 0.7877, TPR: 0.5746

In this test case, the information from the laser scanners is used with the purpose of increase the performance of the system. This required just to include the appropiate nodes for laser data processing, and replace some nodes from the video processing graph to allow data fusion from laser and cameras. An example of vehicle detection based on laser scanners is shown in figure \ref{laser_detection}. This is the output of the node F, points2clusters.

\begin{figure}[htb!]
\centering
\includegraphics[scale=0.4]{fig/4/laser1a.png}
\caption{Vehicle detection using laser scanners}
\label{laser_detection}
\end{figure}

For merging information from laser and cameras, it was decided to perform the fusion at two levels, at the virtual sensor occupancy level and at the flow rate level. Both of these approaches were implemented using average and maximum policies. In the figure \ref{vl_res} is depicted the flow rate output for all of four configurations.

Performing the same frame-based analysis as for the video system, we obtain the confusion matrices presented in table \ref{vl_cf}. It is noted that performing the fusion at virtual sensor occupancy level, the accuracy of the systems has almost no change, 92.2\% and 91.48\% for average and maximum policies respectively. For the TPR, there is a significative increase, having 88.12\% and 63.56\%, also for average and maximum policies.

When the fusion is performed at flow rate level using average policy, the accuracy and TPR values are reduced to 84.45\% and 12.92\%. Using maximum policy, there is a small decrement in the values, with an accuracy of 89.76\% and a TPR of 57.46\%.

\begin{figure}[htb!]
\centering
\includegraphics[scale=0.5]{fig/4/vl_res.pdf}
\caption{Results for the multisensor implementation using four different fusion approaches}
\label{vl_res}
\end{figure}

\begin{table}[htb!]
\footnotesize
\centering
\noindent
\renewcommand\arraystretch{1.5}
\setlength\tabcolsep{0pt}
\begin{tabular}{c c c c c}
  \multirow{10}{*}{\rotatebox{90}{\parbox{1.1cm}{\bfseries\centering Actual\\ value}}} & 
     \multicolumn{4}{c}{\bfseries Outcome (avg vs\_occ)} \\
  & & \bfseries F & \bfseries S & \bfseries total \\
  & F$'$ & \MyBox{1098}{} & \MyBox{148}{} & 1246 \\[2.4em]
  & S$'$ & \MyBox{403}{} & \MyBox{5414}{} & 5817 \\
  & total & 1501 & 5562 &
\end{tabular}

\begin{tabular}{c c c c c}
  \multirow{10}{*}{\rotatebox{90}{\parbox{1.1cm}{\bfseries\centering Actual\\ value}}} & 
     \multicolumn{4}{c}{\bfseries Outcome (max vs\_occ)}  \\
  & & \bfseries F & \bfseries S & \bfseries total \\
  & F$'$ & \MyBox{792}{} & \MyBox{454}{} & 1246 \\[2.4em]
  & S$'$ & \MyBox{148}{} & \MyBox{5669}{} & 5817 \\
  & total & 940 & 6123 &
\end{tabular}

\begin{tabular}{c c c c c}
  \multirow{10}{*}{\rotatebox{90}{\parbox{1.1cm}{\bfseries\centering Actual\\ value}}} & 
     \multicolumn{4}{c}{\bfseries Outcome (avg flow)} \\
  & & \bfseries F & \bfseries S & \bfseries total \\
  & F$'$ & \MyBox{161}{} & \MyBox{1085}{} & 1246 \\[2.4em]
  & S$'$ & \MyBox{13}{} & \MyBox{5804}{} & 5817 \\
  & total & 174 & 6889 &
\end{tabular}

\begin{tabular}{c c c c c}
  \multirow{10}{*}{\rotatebox{90}{\parbox{1.1cm}{\bfseries\centering Actual\\ value}}} & 
     \multicolumn{4}{c}{\bfseries Outcome (max flow)}  \\
  & & \bfseries F & \bfseries S & \bfseries total \\
  & F$'$ & \MyBox{716}{} & \MyBox{530}{} & 1246 \\[2.4em]
  & S$'$ & \MyBox{193}{} & \MyBox{5624}{} & 5817 \\
  & total & 909 & 6154 &
\end{tabular}

\caption{Fusion configurations from top to bottom: Average on virtual sensor occupancy, Maximum on virtual sensor occupancy, Average on flow rate, Maximum on flow rate}
\label{vl_cf}
\end{table}

@@

An intersection management system is a complex set of subsystems, and for a feasible implementation and deployment as a real solution it is needed for it to be easily tested, adapted and enhanced. In this case, after having a well-concieved block design, it was relatively straight forward to start building a video based system for getting an indicator about traffic flow in an intersection using a dataset.

For stakeholders of the system outcome, high level indicators are desired, like traffic status, no matter which low-level hardware or software are used. For this reason, modularity and definitions on data and messages types, allow development of new features without compromising the integrity of the system. Also, new sources of information and data could be easily integrated if the processing of them is made according to the definitions. In that way, it was possible to integrate laser scanner data provided by the dataset, taking into account to generate outputs already defined by the messages specification.

For different sources of information it is not enough to perform a fusion strategy to obtain better results. It is needed to analyse the type of data and the level at which fusion should be performed. Also, it is important to understand if the fusion is competitive, complementary or cooperative, in order to select the appropiate strategy and increase the performance of the system. An example of this, was found when fusing data from camera and lasers at two different levels and using different strategies. It was clear from the results that altough the data was the same, one configuration increased the performace while another one decrease it.
